Lab 2 Final Report: COVID-19 Case Rate vs Population Demographics and Mask Policy

### 2020-12-08 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Reverting changes + new tourism GDP percentage section

The economic standing of a state not only affects the individual households within its population, but may also reflect the economic impact to  their lives as a result of the pandemic. One way that a state may suffer from an increase in the number of COVID cases is from tourists or itinerant workers traveling during the pandemic and spreading the virus. A by-product of this is that residents of the state who work in the tourism industry may be in close contact with the travelers. In addition, if certain key industries in a state, such as tourism, laid off many workers, they would likely have to take lower paying service jobs in the gig economy that force them to be in close contact with others in order to continue to have a livelihood. This would further lead to a higher COVID case rate within the state.

These effects may be captured by a variable that demonstrates how much of a state's GDP is derived from tourism. It would be related to other economic indicator variables, such as the percent of the population living in poverty. While poverty is a broader phenomena, the number of people in poverty in the pandemic could be potentially higher due to negative growth of the tourism industry. Therefore, it is expected that states with a greater focus on tourism may also have more people in poverty in the pandemic. This would take some of the explanatory power of the poverty variable away, resulting in its model coefficient becoming less positive. This would result in a positive omitted variable bias away from zero. Without a clear proxy in the dataset, and with the nature of this variable in touching on the IID assumption between interaction among states, it should be further investigated to understand how much of the original model may be an effect of omitted variable bias.

>>>>

The economic standing of a state not only affects the individual households of its population, but also includes how it interacts with other states and how its peoples’ lives might have been economically changed by the pandemic. One way that a state may increase its number of COVID cases is both from tourists or others visiting during the pandemic and bringing disease, but also if its own population must work in close contact with others serving them in the tourism industry. In addition, if certain key industries to a state, such as tourism, layed off many workers, they would likely have to take lower paying close contact jobs such as in the service or gig economy in order to continue having a livelihood. This would further lead to a higher COVID case rate.

These effects may be included with a variable that demonstrates how much of a state’s GDP is from tourism. It would be related to other economic indicator variables, such as the “real time” poverty rate during the pandemic. Although poverty is a broader phenomena, the number of people in poverty in the pandemic could be higher due to negative growth of the tourism industry. Therefore, it is expected that states with a greater focus on tourism may also have more people in poverty in the pandemic. While poverty rate is included in the model, it is based off of data from 2018 before any pandemic related effects would have occurred. Assuming real time poverty was included in the model, a GDP percentage of tourism would take some of the explanatory power of the poverty variable away, resulting in its model coefficient becoming less positive. This would result in a positive omitted variable bias away from zero. Without a clear proxy in the dataset, and with the nature of this variable in touching on the IID assumption between interaction among states, it should be further investigated to understand a model including only poverty may be an effect of omitted variable bias. Again, because the model in this analysis uses 2018 poverty data, this would not be related to a bias from this omitted variable.
